---
title: "DASC Time Series Forecast Final Project Notebook"
author: "Max Li"
date: "2024-11-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(fable)
library(tsibble)
library(ggplot2)
library(openxlsx)
library(fpp3)
library(patchwork)
```

# Data Preprocessing

## Load the dataset

```{r}
data <- read.csv("aus_mortality.csv")
data <- data %>%
  mutate(Week = yearweek(Week)) %>% 
  as_tsibble(index = Week, key = c(Sex, Age))

head(data)
```

## Split the data into training and testing sets

```{r}
train <- data %>% filter(Week < yearweek("2021 W30"))
test <- data %>% filter(Week >= yearweek("2021 W30"))

glimpse(train)
glimpse(test)
```

## Visualization

```{r}
plot_data <- ggplot(train, aes(x = Week, y = Mortality, color=interaction(Sex, Age))) +
              geom_line() +
              facet_wrap(~Sex + Age, scales = "free_y") +
              theme_minimal() +
              labs(title = "Weekly Mortality Rates by Demographic Group",
                    x = "Week",
                    y = "Mortality Rate")

plot_data +
  theme(
    axis.text.x = element_blank(),  # Remove x-axis labels
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    axis.text.y = element_blank(),  # Remove y-axis labels
    axis.ticks.y = element_blank()   # Remove y-axis ticks
  )
```

# Benchmark Methods

```{r}
# Mean
fit_mean <- train %>%
  model(Mean = MEAN(Mortality))

forecast_mean <- fit_mean %>%
  forecast(new_data = test)

plot_mean <- test %>%
  autoplot(Mortality, color = "grey") +
  autolayer(forecast_mean, level = NULL) +
  labs(
    title = "Mean Method Forecasts by Group",
    x = "Week",
    y = "Mortality"
  ) +
  theme_minimal() +
  facet_wrap(~Sex + Age, scales = "free_y")

plot_mean +
  theme(
    axis.text.x = element_blank(),  # Remove x-axis labels
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    axis.text.y = element_blank(),  # Remove y-axis labels
    axis.ticks.y = element_blank()   # Remove y-axis ticks
  )

# Snaive
fit_naive <- train %>%
  model(Snaive = SNAIVE(Mortality))

forecast_naive <- fit_naive %>%
  forecast(new_data = test)

plot_naive <- test %>%
  autoplot(Mortality, color = "grey") +
  autolayer(forecast_naive, level = NULL) +
  labs(
    title = "Naive Method Forecasts by Group",
    x = "Week",
    y = "Mortality"
  ) +
  theme_minimal() +
  facet_wrap(~Sex + Age, scales = "free_y")

plot_naive +
  theme(
    axis.text.x = element_blank(),  # Remove x-axis labels
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    axis.text.y = element_blank(),  # Remove y-axis labels
    axis.ticks.y = element_blank()   # Remove y-axis ticks
  )

# Drift
fit_drift <- train %>%
  model(Drift = RW(Mortality ~ drift()))

forecast_drift <- fit_drift %>%
  forecast(new_data = test)

plot_drift <- test %>%
  autoplot(Mortality, color = "grey") +
  autolayer(forecast_drift, level = NULL) +
  labs(
    title = "Drift Method Forecasts by Group",
    x = "Week",
    y = "Mortality"
  ) +
  theme_minimal() +
  facet_wrap(~Sex + Age, scales = "free_y")

plot_drift +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

# Time Series Regression Model

```{r}
tslm_model <- train %>%
  model(TSLM_1 = TSLM(Mortality ~ trend() + season()))

forecast_tslm <- tslm_model %>%
  forecast(new_data = test)

plot_tslm <- test %>%
  autoplot(Mortality, color = "grey") +
  autolayer(forecast_tslm, level = NULL, aes(color = interaction(Sex, Age))) +
  labs(
    title = "TSLM Method Forecasts by Group",
    x = "Week",
    y = "Mortality"
  ) +
  theme_minimal() +
  facet_wrap(~Sex + Age, scales = "free_y")

plot_tslm +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

## Ljung-Box Test

```{r}
augment(tslm_model) %>%
  features(.innov, ljung_box, lag = 104)
```

# Auto ARIMA Model

```{r}
auto_arima <- train %>%
  model(ARIMA = ARIMA(Mortality))

forecast_arima <- auto_arima %>%
  forecast(new_data = test)

plot_arima <- test %>%
  autoplot(Mortality, color = "grey") +
  autolayer(forecast_arima, level = NULL, aes(color = interaction(Sex, Age))) +
  labs(
    title = "ARIMA Model Forecasts Faceted by Groups",
    x = "Week",
    y = "Mortality"
  ) +
  theme_minimal() +
  facet_wrap(~Sex + Age, scales = "free_y")

plot_arima +
  theme(
    axis.text.x = element_blank(),  # Remove x-axis labels
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    axis.text.y = element_blank(),  # Remove y-axis labels
    axis.ticks.y = element_blank()   # Remove y-axis ticks
  )
```

## Ljung-Box Test

```{r}
augment(auto_arima) %>%
  features(.innov, ljung_box, lag = 104)
```

# Cross Validation

## CV training set split
```{r}
cv_splits <- train %>%
  stretch_tsibble(.init = 104, .step = 52) %>%
  select(Week, Sex, Age, Deaths, Mortality)
```
## CV process
```{r}
cv_results <- cv_splits %>%
  group_by(Sex, Age) %>%
  model(
    MEAN = MEAN(Mortality),
    SNAIVE = SNAIVE(Mortality),
    DRIFT = RW(Mortality ~ drift()),
    TSLM = TSLM(Mortality ~ trend() + season()),
    ARIMA = ARIMA(Mortality),
  ) %>%
  forecast(h = "1 year") %>%
  accuracy(train)
```

## Save CV results

```{r}
cv_summary <- cv_results %>%
  group_by(.model) %>%
  summarise(AVG_MAE = mean(MAE), AVG_RMSE = mean(RMSE), AVG_MAPE = mean(MAPE))

write_csv(cv_summary, "cv_summary.csv")
```

# Accuracy Results

## Extract accuracy results
```{r}
accuracy_mean <- forecast_mean %>% accuracy(test)
accuracy_naive <- forecast_naive %>% accuracy(test)
accuracy_drift <- forecast_drift %>% accuracy(test)
accuracy_tslm <- forecast_tslm %>% accuracy(test)
accuracy_arima <- forecast_arima %>% accuracy(test)
```

## Merge result by rows
```{r}
model_comparison <- bind_rows(
  accuracy_mean %>% mutate(Model = "Mean"),
  accuracy_naive %>% mutate(Model = "Naive"),
  accuracy_drift %>% mutate(Model = "Drift"),
  accuracy_tslm %>% mutate(Model = "TSLM"),
  accuracy_arima %>% mutate(Model = "ARIMA"),
)
```

## Save results
```{r}
accuracy_summary <- model_comparison %>%
  group_by(.model) %>%
  summarise(AVG_MAE = mean(MAE), AVG_RMSE = mean(RMSE), AVG_MAPE = mean(MAPE))

write_csv(accuracy_summary, "accuracy_summary.csv")
```

## Result Visualization

```{r}
accuracy <- read.csv("accuracy_summary.csv")

# Plot for MAE
plot_mae <- ggplot(accuracy, aes(x = Model, y = Average.MAE, fill = Model)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average MAE",
    x = "Model",
    y = " "
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Plot for RMSE
plot_rmse <- ggplot(accuracy, aes(x = Model, y = Average.RMSE, fill = Model)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average RMSE",
    x = "Model",
    y = " "
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Plot for MAPE
plot_mape <- ggplot(accuracy, aes(x = Model, y = Average.MAPE, fill = Model)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average MAPE",
    x = "Model",
    y = " "
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

combined_plot <- plot_mae + plot_rmse + plot_mape + plot_layout(nrow = 1)

combined_plot
```
